Ground state energy of quantum dots using the coupled cluster method

### 2018-04-30 project_2/article/article.tex
Add basis figure

        \subsubsection{Harmonic oscillator basis}
            We will be looking at a system of two-dimensional quantum dots with
            a Coulomb repulsion.  If we assume, or make it so, that the
            repulsive two-body part is small we can use the harmonic oscillator
            functions as our basis functions and compute the matrix elements
            before feeding it into the coupled cluster code. We will label these
            basis functions using greek letters, viz.
            \begin{align}
                h\ket{\alpha} = \epsilon_{\alpha}\ket{\alpha}.
            \end{align}
            In polar coordinates we can write the harmonic oscillator
            wavefunction for a single particle in two dimensions\footnote{Note
            that this is without spin. As we are looking at fermions this means
            that each mode of the harmonic oscillator functions will be repeated
            twice.}.
            \begin{align}
                \phi_{nm}(r, \theta)
                &=
                N_{nm}
                %a\sqrt{\frac{n!}{\pi(n + \abs{m})!}}
                (ar)^{\abs{m}}L_n^{\abs{m}}(a^2 r^2)e^{-a^2 r^2/2}
                e^{im\theta},
            \end{align}
            where $a = \sqrt{m\omega/\hbar}$ is the Bohr radius, $L_n^{\abs{m}}$
            is the associated Laguerre polynomials, $n$ and $m$ are the
            principal and azimuthal quantum numbers respectively and $N_{nm}$ is
            a normalization constant given by
            \begin{align}
                N_{nm}
                &= a\sqrt{\frac{n!}{\pi(n + \abs{m})!}}.
            \end{align}

>>>>

        \subsubsection{Harmonic oscillator basis}
            We will be looking at a system of two-dimensional quantum dots with
            a Coulomb repulsion.  If we assume, or make it so, that the
            repulsive two-body part is small we can use eigenfunctions of the
            one-body part our basis. In this case we have two-dimensional
            harmonic oscillator functions as eigenfunctions. We can then compute
            the matrix elements, $h_q^p$ and $u_{rs}^{pq}$, before feeding these
            into the coupled cluster code.

            In polar coordinates we can write the harmonic oscillator
            wavefunction for a single particle in two dimensions
            as\footnote{Note that this is without spin. As we are looking at
            fermions this means that each mode of the harmonic oscillator
            functions will be repeated twice.}.
            \begin{align}
                \phi_{nm}(r, \theta)
                &=
                N_{nm}
                (ar)^{\abs{m}}L_n^{\abs{m}}(a^2 r^2)e^{-a^2 r^2/2}
                e^{im\theta},
            \end{align}
            where $a = \sqrt{m\omega/\hbar}$ is the Bohr radius, $L_n^{\abs{m}}$
            is the associated Laguerre polynomials, $n$ and $m$ are the
            principal and azimuthal quantum numbers respectively and $N_{nm}$ is
            a normalization constant given by
            \begin{align}
                N_{nm}
                &= a\sqrt{\frac{n!}{\pi(n + \abs{m})!}}.
            \end{align}
            Included is also the spin, $\sigma$, of the wavefunction, which can
            be either up or down. This means each level, $(n, m)$, is doubly
            occupied. We also have that the wavefunctions are orthonormal
            \begin{align}
                \braket{n_1m_1\sigma_1}{n_2m_2\sigma_2}
                = \delta_{n_1n_2}\delta_{m_1m_2}\delta_{\sigma_1\sigma_2}.
            \end{align}
            The eigenenergy of a single harmonic oscillator is given by
            \begin{align}
                \epsilon_{nm} &= \hbar\omega(2n + \abs{m} + 1).
            \end{align}
            As the wavefunctions are eigenfunctions to the one-body part we will
            get a diagonal one-body matrix $h$,
            \begin{align}
                h\ket{\alpha} = \epsilon_{\alpha}\ket{\alpha},
            \end{align}
            where we have labeled the harmonic oscillator basis functions with
            greek letters, i.e., $\alpha = (n_{\alpha}, m_{\alpha},
            \sigma_{\alpha})$. Our next job is now to create a mapping between
            the quantum number $\alpha$ and the three quantum numbers $n$, $m$
            and $\sigma$.

            \begin{figure}
                \begin{center}
                    \begin{tikzpicture}[scale=0.8]
                        \begin{scope}
                            \foreach \i in {1, 2, 3} {
                                \draw(-1, \i - 1) node[anchor=east]
                                {$\epsilon = \i$};
                            }

                            % Highest energy level
                            \foreach \i in {0, 3, 6} {
                                \draw (\i, 2) -- (\i + 2, 2);
                                \node at (\i + 0.75, 2) {$\uparrow$};
                                \node at (\i + 1.25, 2) {$\downarrow$};
                            }
                            \node[below, inner sep=.2cm] at (1, 2)
                            {$n = 0, m = -2$};
                            \node[below, inner sep=.2cm] at (4, 2)
                            {$n = 1, m = 0$};
                            \node[below, inner sep=.2cm] at (7, 2)
                            {$n = 0, m = 2$};

                            % Middle energy level
                            \foreach \i in {1.5, 4.5} {
                                \draw (\i, 1) -- (\i + 2, 1);
                                \node at (\i + 0.75, 1) {$\uparrow$};
                                \node at (\i + 1.25, 1) {$\downarrow$};
                            }
                            \node[below, inner sep=.2cm] at (2.5, 1)
                            {$n = 0, m = -1$};
                            \node[below, inner sep=.2cm] at (5.5, 1)
                            {$n = 0, m = 1$};

                            % Lowest energy level
                            \draw (3, 0) -- (5, 0);
                            \node at (3 + 0.75, 0) {$\uparrow$};
                            \node at (3 + 1.25, 0) {$\downarrow$};
                            \node[below, inner sep=.2cm] at (4, 0)
                            {$n = 0, m = 0$};
                        \end{scope}
                    \end{tikzpicture}
                \end{center}
                \caption{In this plot we can see the energy degeneracy of the
                lowest three energy levels in the two-dimensional quantum dot.
                Each arrow representes a spin up or a spin down state with the
                quantum numbers $n$ and $m$ as listed below.}
                \label{fig:basis_states}
            \end{figure}
